Sir Alex Ferguson's Continued Tenure Could Have Only Marginal Improvements over Manchester United's EPL Performance

### 2024-04-16 paper/paper.qmd
Added Abstract. Editted title. Added sketches. Used styler to style code. Deleted files not used.

---
title: "Sir Alex Ferguson's Continued Tenure Could Have Sustained Manchester United's EPL Dominance"
subtitle: "An Analysis of an Increase in League Points"
author: 
  - Aryaman Sharma
thanks: "Code and data are available at: https://github.com/Ary4m3n/manchester-united.git"
date: today
date-format: long
abstract: "This study investigates the unexpected surge in death rates during the COVID-19 pandemic in Russia, a country considered well-prepared according to the Global Health Security (GHS) Index. Building upon Nuzzo and Ledesma's work, we analyze the impact of COVID-19 in Russia, incorporating factors such as the GHS Index and excess death rates. Our findings reveal a substantial increase in excess death rates in Russia, diverging from the lower rates observed in nations deemed well-prepared. This research contributes to our understanding of the repercussions of the COVID-19 outbreak in Russia, highlighting areas for improvement in prepared countries to mitigate the effects of future pandemics."
format: 
  pdf:
    tbl-pos: "h"
    fig-pos: "h"
toc: true
number-sections: true
bibliography: references.bib
---

@@

The aim of this paper is twofold, where in the paper will first analyze the differences in performance between 2004-2013 (Sir Alex's era) and 2014-2023 (Post Sir Alex's era). Then, we will use this knowledge to build two Bayesian Models to predict and further analyze the estimand: Average League Points in each season for 2004-2013 and 2014-2023. The paper will also illustrate *how much* better Manchester United's position would have been in 2014-2023 if Sir Alex Ferguson would not have retired, helping us contribute positively to the debate about him being the greatest manager of all time.

@@

The datasets under examination in this paper were obtained from the R package and API `worldfootballR` [@citeWorldfootballR]. Employing the open-source `R` programming language [@citeR], we conducted the cleaning and analysis procedures, leveraging several R libraries and packages such as `tidyverse` [@citeTidyverse], `janitor` [@citeJanitor], `ggplot2` [@citeGgplot], `knitr` [@citeKnitr], `readr` [@citeReadr], `dplyr` [@citeDplyr] and `arrow` [@citeArrow]. We used `rstanarm` [@rstanarm] for the models and `modelsummary` [@citeModelsummary] for the model summary.

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl-model-predictions
#| tbl-cap: "Comparision of Predicted League Points for Sir Alex's Continued Era and Other 8 Mangers with Actual League Points (2014-2023)"
#| warning: false

@@

@fig-prediction shows a season by season comparision of how well Sir Alex would have done in terms of league points, with how well the other 8 managers are predicted to do, and how well the team did in real. It is important to note here that looking at @tbl-model-predictions and @fig-prediction, we realise that although Sir Alex is seen to would have done better on average between 2014 and 2023, the difference is not as much as we expected. For 8 out of 10 seasons, Sir Alex is predicted to have done better than the team performed. But as we will observe in @fig-prediction-delta and @tbl-prediction-delta, the difference is not a lot.

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: tbl-prediction-delta
#| tbl-cap: "Comparision of Differences between League Points for Sir Alex's Continued Era and Other 8 Mangers and between Actual League Points (2014-2023)"

@@

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-prediction-delta
#| fig-cap: "Plot for comparision of differences between League Points for Sir Alex's Continued Era and Other 8 Mangers and between Actual League Points (2014-2023)"
#| fig-width: 8

>>>>

---
title: "Sir Alex Ferguson's Continued Tenure Could Have Only Marginal Improvements over Manchester United's EPL Performance"
subtitle: "An Analysis of an Increase in League Points"
author: 
  - Aryaman Sharma
thanks: "Code and data are available at: https://github.com/Ary4m3n/manchester-united.git"
date: today
date-format: long
abstract: "Manchester United has been a historical football club. This paper examines Manchester United's performance from 2004-2013 under Sir Alex Ferguson's golden era and compares it with the period from 2014-2023, following his retirement. Using a Bayesian analysis model, this paper attempts to predict potential league points for if Sir Alex would have continued managing (2014-2023). Results predict only a marginal improvement in club's performance. This reasearch delves into recent increase in competition and other factor as a possible explanation."
format: 
  pdf:
    tbl-pos: "h"
    fig-pos: "h"
toc: true
number-sections: true
bibliography: references.bib
---

@@

The aim of this paper is twofold, where in the paper will first analyze the differences in performance between 2004-2013 (Sir Alex's era) and 2014-2023 (Post Sir Alex's era). Then, we will use this knowledge to build two Bayesian Models to predict and further analyze the estimand: Average League Points in each season for 2004-2013 and 2014-2023. The paper will also illustrate *how much* better Manchester United's position would have been in 2014-2023 if Sir Alex Ferguson would not have retired, helping us contribute to the debate about the greatest manager of all time.

@@

The datasets under examination in this paper were obtained from the R package and API `worldfootballR` [@citeWorldfootballR]. Employing the open-source `R` programming language [@citeR], we conducted the cleaning and analysis procedures, leveraging several R libraries and packages such as `tidyverse` [@citeTidyverse], `janitor` [@citeJanitor], `ggplot2` [@citeGgplot], `knitr` [@citeKnitr], `readr` [@citeReadr], `dplyr` [@citeDplyr], `styler` [@citeStyler] and `arrow` [@citeArrow]. We used `rstanarm` [@rstanarm] for the models and `modelsummary` [@citeModelsummary] for the model summary.

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl-model-predictions
#| tbl-cap: "Comparison of Predicted League Points for Sir Alex's Continued Era and Other 8 Mangers with Actual League Points (2014-2023)"
#| warning: false

@@

@fig-prediction shows a season by season comparison of how well Sir Alex would have done in terms of league points, with how well the other 8 managers are predicted to do, and how well the team did in real. It is important to note here that looking at @tbl-model-predictions and @fig-prediction, we realise that although Sir Alex is seen to would have done better on average between 2014 and 2023, the difference is not as much as we expected. For 8 out of 10 seasons, Sir Alex is predicted to have done better than the team performed. But as we will observe in @fig-prediction-delta and @tbl-prediction-delta, the difference is not a lot.

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: tbl-prediction-delta
#| tbl-cap: "Comparison of Differences between League Points for Sir Alex's Continued Era and Other 8 Mangers and between Actual League Points (2014-2023)"

@@

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-prediction-delta
#| fig-cap: "Plot for comparison of differences between League Points for Sir Alex's Continued Era and Other 8 Mangers and between Actual League Points (2014-2023)"
#| fig-width: 8
